---
title: "Day 5 Notes: Functions"
author: "Erica Bishop"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Testing Functions

-   informal testing in the console doesn't leave a record, so it becomes easy to break code that used to work
-   to get around this, do a **unit test**

**Unit Tests**
-   in R, use a package called "testthat"
-   testthat is supposed to make testing easy and "addictive"
-   describe what you want a function to do
Examples of unit tests build into testthat:
-   expect_equal() does the function return expected values?
-   expect_type() does the function return the expected type?
-   expect_length() if a vector is returned, is it the expected length?
-   expect_true() does the function return TRUE?

If the unit test passes, nothing will happen
If it **fails** then you'll get an error message


### Packages

How do we find, vet, and use useful packages?

Follow the checklist, especially if its not on CRAN or rOpenSci:
-   look at list of updates on CRAN - how frequently / recently updated? 
-   is it under active development? (might not want to use an immature package where things are still changing) but also active developers can be responsive to bugs and fixes
-   recently updated
-   documentation available - should have substantial documentation
-   is it used and published in similar projects? (precedent set) (google it)
-   Liscensing?
-   semantics intuitive? (will the packages coding style / preferred syntax be easy for you to use?)
-   functionality? does it do what you need it?
-   arguments support your needs?
-   what dependencies does the package have and are they reasonable? 

Can also look at the package on github

## Citing packages

R in the roxygen template has a space for how to cite the package. can alse use `citation("package")` to find the citation


## how to recase a character as numeric

```{r}

penguins_yr_chr <- penguins |> 
  mutate(year = as.character(year))

penguins_year_num <- penguins_yr_chr |> 
  mutate(year = as.numeric(year))

```

Another quick example on subsetting:

```{r}

penguins_ad_ch <- penguins_year_num |> 
  filter(species %in% c("Adelie", "Chinstrap"))

```

the %in% operator looks in each row to see if the variable in that row exists in the vector specified.
this is VERY DIFFERENT from using ==


